### 2015-04-17 implementation-specs.tex
add specialization.

So far, we have shown examples of specs being used to request constraints from the
command line, when {\tt spack install} is invoked.  However, the user is not the only
source of constraints.  Applications may require specific versions of dependencies,
and it is often desirable to write constraints like this into a package file.  For
example, the ROSE compiler only builds with a certain version of the {\tt boost} library.
At first glance, the {\tt depends\_on()} directive in Figure~\ref{fig:mpileaks} look
like they take simple package names.  However, a package name is also a spec, and
the same constraint syntax usable from the command line can be applied inside directives.
So, the ROSE compiler can simply write:
%
\begin{lstlisting}
depends_on(`boost@1.54.0')
\end{lstlisting}
%
This constraint will be incorporated into the initial DAG node generated from
the ROSE package.

Spack also accepts spec syntax for conditionally patching packages.
Spack offers a {\tt patch} directive that uses specs as predicates in an
optional {\tt when} parameter.
These concise directives in the Python package ensure that very specific patches
are only applied to the Python source code when it is being built on Blue Gene/Q,
with the appropriate compiler:
%
\begin{lstlisting}
patch(`python-bgq-xlc.patch',   when=`=bgq%xlc')
patch(`python-bgq-clang.patch', when=`=bgq%clang')
\end{lstlisting}
%
The {\tt when} clause is matched against the Python package's spec
to determine whether the patch will be applied.

Constraint tests can also be used directly within package source code.
For example, in the {\tt install} method, packages can test the versions of dependencies
using constraint syntax:
%
\begin{lstlisting}
  if '' in spec:
  elif '
  '
\end{lstlisting}
%

@@

In some cases, simple conditional logic

In our experience maintaining packages over time, we have found tha

>>>>

So far, we have shown examples of specs being used to request constraints from the
command line, when {\tt spack install} is invoked.  However, the user is not the only
source of constraints.  Applications may require specific versions of dependencies,
and it is often desirable to write constraints like this into a package file.  For
example, the ROSE compiler only builds with a certain version of the {\tt boost} library.
At first glance, the {\tt depends\_on()} directive in Figure~\ref{fig:mpileaks} look
like they take simple package names.  However, a package name is also a spec, and
the same constraint syntax usable from the command line can be applied inside directives.
So, the ROSE compiler can simply write:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('boost@1.54.0')
\end{minted}
%
This constraint will be incorporated into the initial DAG node generated from
the ROSE package.

Spack's {\tt patch} directive also accepts spec syntax as a predicates in an
optional {\tt when} parameter.  For example, these concise directives in the 
Python package ensure that specific patches are applied to the Python source
code when it is being built on Blue Gene/Q, with the appropriate compiler:
%
\begin{minted}[fontsize=\scriptsize]{python}
    patch('python-bgq-xlc.patch',   when='=bgq%xlc')
    patch('python-bgq-clang.patch', when='=bgq%clang')
\end{minted}
%
Constraints in the {\tt when} clause are matched against the Python package spec.
Outside of directives, constraints can be used directly
with the {\tt spec} object in the {\tt install} method:
%
\begin{minted}[fontsize=\scriptsize]{python}
  def install(self, spec, prefix):
      if spec.satisfies('%gcc'):
          # Handle gcc
      elif spec.satisfies('%xlc'):
          # Handle XL compilers
      ...
\end{minted}
%

@@

In our experience maintaining packages at LLNL, we have sometimes had to change
entire package build scripts due to large changes in the way certain packages build.
This can be very cumbersome, and it is difficult to maintain both the old and new
version of a build script, but we must if we want to keep installing older versions
for users who rely on them.

\begin{figure}
\begin{minted}[linenos,
               numbersep=5pt,
			   fontsize=\scriptsize,
               frame=lines,
               framesep=2mm]{python}
class Dyninst(Package):
    """"Dyninst binary instrumentation and analysis tool."""
    url="http://www.paradyn.org/release8.1/DyninstAPI-8.1.1.tgz"
    version('8.1.1', 'd1a04e995b7aa70960cd1d1fac8bd6ac')
    depends_on("libelf")                                                                            
    depends_on("libdwarf")                                                                          
    depends_on("boost@1.42:")

    # default build version uses cmake                                                                        
    def install(self, spec, prefix):
        with working_dir('spack-build', create=True):                                               
            cmake('..', *std_cmake_args)
            make()                                                                                  
            make("install")

    # when version is 8.1 or lower, use configure.                                                                                                    
    @when('@:8.1')                                                                                  
    def install(self, spec, prefix):                                                                
        configure("--prefix=" + prefix)                                                             
        make()                                                                                      
        make("install")     
\end{minted}
\caption{
    Specialized {\tt install} method in Dyninst. 
	\label{fig:specialization}
}
\end{figure}

For cases like this, Spack provides functionality that allows Python functions to have
multiple definitions, with some specialized for particular configurations of the package.
This allows us to have two separate implementations of {\tt install} or {\it any} method 
in a package class. Figure~\ref{fig:specialization} shows how this is used in the Dyninst
package.  The {\tt @when} directive is a Python decorator: a higher order function that 
takes a function definition as a parameter and returns a new function to put in its place.
Here, we replace the function with a callable multi-function dispatch object, and we 
integrate the predicate check into the function dispatch mechanism.  Here, the
{\tt when} condition is true when Dyninst is at version 8.1 or lower, and in those cases
the package will use the {\tt configure}-based build.  By default if no predicate matches,
{\tt install} will simply use the default CMake-based implementation. The simple {\tt @when}
annotation allows us to maintain our old build code alongside the new version without 
accumulating complex logic in a single {\tt install} function.
